Rust provides three distribution channels for Rust: nightly, beta, and stable. Unstable features are only available 
on nightly Rust. For more details on this process, see \href{http://blog.rust-lang.org/2014/10/30/Stability.html}{Stability 
as a deliverable}.

\blank

To install nightly Rust, you can use \code{rustup.sh}:

\begin{verbatim}
$ curl -s https://static.rust-lang.org/rustup.sh | sh -s -- --channel=nightly
\end{verbatim}

If you're concerned about the \href{http://curlpipesh.tumblr.com/}{potential insecurity} of using \code{curl | sh}, 
please keep reading and see our disclaimer below. And feel free to use a two-step version of the installation and examine 
our installation script:

\begin{verbatim}
$ curl -f -L https://static.rust-lang.org/rustup.sh -O
$ sh rustup.sh --channel=nightly
\end{verbatim}

If you're on Windows, please download either the 
\href{https://static.rust-lang.org/dist/rust-nightly-i686-pc-windows-gnu.msi}{32-bit installer} or the 
\href{https://static.rust-lang.org/dist/rust-nightly-x86_64-pc-windows-gnu.msi}{64-bit installer} and run it.

\subsection*{Uninstalling}

If you decide you don't want Rust anymore, we'll be a bit sad, but that's okay. Not every programming language is great 
for everyone. Just run the uninstall script:

\begin{verbatim}
$ sudo /usr/local/lib/rustlib/uninstall.sh
\end{verbatim}

If you used the Windows installer, re-run the \code{.msi} and it will give you an uninstall option.

\blank

Some people, and somewhat rightfully so, get very upset when we tell you to \code{curl | sh}. Basically, when you do this, 
you are trusting that the good people who maintain Rust aren't going to hack your computer and do bad things. That's a good 
instinct! If you're one of those people, please check out the documentation on 
\href{https://github.com/rust-lang/rust\#building-from-source}{building Rust from Source}, or 
\href{https://www.rust-lang.org/install.html}{the official binary downloads}.

\blank

Oh, we should also mention the officially supported platforms:

\begin{itemize}
  \item{Windows (7, 8, Server 2008 R2)}
  \item{Linux (2.6.18 or later, various distributions), x86 and x86-64}
  \item{OSX 10.7 (Lion) or greater, x86 and x86-64}
\end{itemize}

We extensively test Rust on these platforms, and a few others, too, like Android. But these are the ones most likely 
to work, as they have the most testing.

\blank

Finally, a comment about Windows. Rust considers Windows to be a first-class platform upon release, but if we're honest, 
the Windows experience isn't as integrated as the Linux/OS X experience is. We're working on it! If anything does not work, 
it is a bug. Please let us know if that happens. Each and every commit is tested against Windows like any other platform.

\blank

If you've got Rust installed, you can open up a shell, and type this:

\begin{verbatim}
$ rustc --version
\end{verbatim}

You should see the version number, commit hash, commit date and build date:

\begin{verbatim}
rustc 1.0.0-nightly (f11f3e7ba 2015-01-04) (built 2015-01-06)
\end{verbatim}

If you did, Rust has been installed successfully! Congrats!

\blank

This installer also installs a copy of the documentation locally, so you can read it offline. On UNIX systems, 
\code{/usr/local/share/doc/rust} is the location. On Windows, it's in a \code{share/doc} directory, inside wherever you 
installed Rust to.

\blank

If not, there are a number of places where you can get help. The easiest is the 
\href{irc://irc.mozilla.org/\#rust}{\#rust IRC channel on irc.mozilla.org}, which you can access through 
\href{http://chat.mibbit.com/?server=irc.mozilla.org\&channel=\%23rust}{Mibbit}. Click that link, and you'll be chatting with 
other Rustaceans (a silly nickname we call ourselves), and we can help you out. Other great resources include the 
\href{https://users.rust-lang.org/}{user's forum}, and \href{http://stackoverflow.com/questions/tagged/rust}{Stack Overflow}.

\section{Compiler Plugins}
\label{sec:nightly_compilerPlugins}
\input{src/nightly_rust/compiler_plugins.tex}

\section{Inline Assembly}
\label{sec:nightly_inlineAssembly}
\input{src/nightly_rust/inline_assembly.tex}

\section{No stdlib}
\label{sec:nightly_nostdlib}
\input{src/nightly_rust/no_stdlib.tex}

\section{Intrinsics}
\label{sec:nightly_intrinsics}
\input{src/nightly_rust/intrinsics.tex}

\section{Lang items}
\label{sec:nightly_langItems}
\input{src/nightly_rust/lang_items.tex}

\section{Advanced Linking}
\label{sec:nightly_advancedLinking}
\input{src/nightly_rust/advanced_linking.tex}

\section{Benchmark Tests}
\label{sec:nightly_benchmarkTests}
\input{src/nightly_rust/benchmark_tests.tex}

\section{Box Syntax and Patterns}
\label{sec:nightly_boxSyntaxAndPatterns}
\input{src/nightly_rust/box_syntax_and_patterns.tex}

\section{Slice patterns}
\label{sec:nightly_slicePatterns}
\input{src/nightly_rust/slice_patterns.tex}

\section{Associated Constants}
\label{sec:nightly_associatedConstants}
\input{src/nightly_rust/associated_constants.tex}

\section{Custom Allocators}
\label{sec:nightly_customAllocators}
\input{src/nightly_rust/custom_allocators.tex}
